---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

**Authors:** [Gede Primahadi Wijaya Rajeg](https://figshare.com/authors/Gede_Primahadi_Wijaya_Rajeg/1234749)<br/>
**License:** [GPL-2](https://www.gnu.org/licenses/old-licenses/gpl-2.0.en.html)<br/>

[![Travis-CI Build Status](https://travis-ci.org/gederajeg/collogetr.svg?branch=master)](https://travis-ci.org/gederajeg/collogetr)<br/>
[![Coverage Status](https://img.shields.io/codecov/c/github/gederajeg/collogetr/master.svg)](https://codecov.io/github/gederajeg/collogetr?branch=master)


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
scaps <- function(concept) {
  paste("<span style='font-variant:small-caps;'>", concept, "</span>", sep="")
}
```

# collogetr

## Overview

`collogetr` has one function (viz. `colloc_leipzig()`) to retrieve window-span collocates for a set of word forms (viz. the *node word*) from the (Indonesian) [Leipzig Corpora](http://wortschatz.uni-leipzig.de/en/download). There are two functions to process the output of `colloc_leipzig()` into tabular formats as input for **association measure** between the collocates and the node, as in the [collostructional/collocation analysis](http://www.linguistics.ucsb.edu/faculty/stgries/teaching/groningen/index.html). These functions are `assoc_prepare()` and `assoc_prepare_dca()`. The former generates input for [*Simple Collexeme/Collocational Analysis*](http://www.linguistics.ucsb.edu/faculty/stgries/research/2003_AS-STG_Collostructions_IJCL.pdf) computed using `collex_fye()`, meanwhile the latter uses the output of `assoc_prepare()` to generate input for [*Distinctive Collexeme/Collocates Analysis*](http://www.linguistics.ucsb.edu/faculty/stgries/research/2004_STG-AS_ExtendingCollostructions_IJCL.pdf) computed using `collex_fye_dca()`. The package is built on top of the core [tidyverse](https://www.tidyverse.org) packages.

## Installation

Install from GitHub with [devtools](https://github.com/hadley/devtools):

```{r eval = FALSE}
library(devtools)
install_github("gederajeg/collogetr")
```

## Usages

### Load `collogetr`
```{r load-collogetr, message = FALSE}
library(collogetr)
```

### Citation `collogetr`

To cite the package in publication, type as follows:
```{r cite-collogetr, eval = TRUE}
citation("collogetr")
```

### Package data

The package has three data sets for demonstration. The important one is the `demo_corpus_leipzig` whose documentation can be accessed via `?demo_corpus_leipzig`. Another data is a list of Indonesian stopwords (i.e. `stopwords`) that can be filtered out when performing collocational measure. The last one is `leipzig_corpus_path` containing character vector of full path to my Leipzig Corpus files in my computer.

#### Accepted inputs

`colloc_leipzig()` accepts two types of input data:

1. A named-list object with character-vector elements of each Leipzig Corpus Files, represented by `demo_corpus_leipzig` and the format of which is shown below:

```{r leipzig_list_input}
lapply(demo_corpus_leipzig[1:2], sample, 2)
```

2. Full-paths to the Leipzig Corpus plain texts, as in the `leipzig_corpus_path`.

```{r leipzig_path_input}
leipzig_corpus_path[1:2]
```

### Demo

#### Retrieving the collocates

Here is how to retrieve the collocates for the Indonesian verb *mengatakan* 'to say sth.'. The function will print out progress messages of the stages onto the console. It generates warning(s) when a search pattern or node word is not found in a corpus file or in all loaded corpus files.

```{r example, eval = TRUE, message = FALSE, warning = FALSE}
out <- colloc_leipzig(leipzig_corpus_list = demo_corpus_leipzig,
                       pattern = "mengatakan",
                       window = "r",
                       span = 1L,
                       save_interim = FALSE)
```

In the example above, the collocates are restricted to those occurring _one_ word (i.e. `span = 1L`) to the _right_ (`window = "r"`) of *mengatakan* 'to say'. The `"r"` character in `window` stands for *right*-side collocates (`"l"` for *left*-side collocates and `"b"` for *both* right- and left-side collocates). The `span` argument requires integer (i.e., a whole number) to indicate the range of words covered in the specified window. The `pattern` argument requires one or more exact word forms; if more than one, put into a character vector (e.g., `c("mengatakan", "menjanjikan")`). The `pattern` argument also accepts strings marked with *word boundary* character (i.e. `"\\b"`), thus `c("\\bmengatakan\\b", "\\bmenjanjikan\\b")`. 

The `save_interim` is `FALSE` means that no output is saved into the computer, but in the console (i.e., in the `out` object). If `save_interim = TRUE`, the function will save the outputs into the files in the computer. `colloc_leipzig()` has specified the default file names for the outputs via these arguments: (i) `freqlist_output_file`, (ii) `colloc_output_file`, (iii) `corpussize_output_file`, and (iv) `search_pattern_output_file`. It is recommended that the output filenames are stored as a character vector. See **Examples** "(2)" in the documentation of `colloc_leipzig()` for a call when `save_interim = TRUE`.

#### Exploring the output of `colloc_leipzig()`.

The output of `colloc_leipzig()` is a list of `r length(out)` elements:

1. `colloc_df`; a table/tibble of raw collocates data with columns for:
    1. corpus names
    2. sentence id in which the collocates and the node word(s) are found
    3. the collocates (column `w`)
    5. the span information (e.g., `"r1"` for one-word, right-side collocates)
    6. the node word
    7. the text/sentence match in which the collocates and the node are found
2. `freqlist_df`; a table/tibble of word-frequency list in the loaded corpus
3. `corpussize_df`; a table/tibble of total word-tokens in the loaded corpus
4. `pattern`; a character vector of the search pattern/node

```{r colloc_output}
str(out)
```

The `freqlist_df` and `corpussize_df` are important for performing the collocational strength measure for the search pattern with the collocates.

#### Preparing input data for *Simple Collexeme/Collocational Analysis* (SCA).

First we need to call `assoc_prepare()` for generating the data for SCA with `collex_fye()`. The demo illustrates it with on-console output of `colloc_leipzig()`. Cf. the **Examples** "2.2" in the documentation for `assoc_prepare()` for handling saved outputs (`?assoc_prepare()`).

```{r assoc-prepare}
assoc_tb <- assoc_prepare(colloc_out = out, 
                          window_span = "r1",
                          per_corpus = FALSE, # combine all data across corpus
                          stopword_list = collogetr::stopwords,
                          float_digits = 3L)
```

Inspect the output of `assoc_prepare()`
```{r assoc-prepare-head}
head(assoc_tb)
```

The `assoc_prepare()` and `collex_fye()` functions are designed following the tidy principle so that the association/collocation measure is performed in a row-wise fashion, benefiting from the combination of [*nested* column](http://r4ds.had.co.nz/many-models.html#list-columns-1) for the input-data (using `tidyr::nest()`) and `purrr`'s `map_*` function. `assoc_prepare()` includes calculating the expected co-occurrence frequencies between the collocates/collexemes and the node word/construction. 

The column `data` in `assoc_tb` above consists of nested tibble/table as a list. Each contains required data for performing association measure for each of the collocates in column `w`. This nested column can be inspected as follows (for the first row, namely for the word *pihaknya* 'the party').

```{r inspect-nested-column}
# get the tibble in the `data` column for the first row
assoc_tb$data[[1]]
```

Column `a` indicates the co-occurrence frequency between the node word and the collocates column `w`, meanwhile `a_exp` indicates the *expected co-occurrence frequency* between them. The `n_w_in_corp` represents the total token/occurrence frequency of a given collocate. The `n_pattern` stores the total token/occurrence frequency of the node word in the corpus. Column `b`, `c`, and `d` are required for the association measure that is essentially based on 2-by-2 crosstabulation table. The `assoc` column indicates whether the value in `a` is higher than that in `a_exp`, thus indicating *attraction* or *positive association* between the node word and the collocate. The reverse is *repulsion* or *negative association* when the value in `a` is less/lower than that in `a_exp`.

#### *Simple Collexeme/Collocates Analysis (SCA)*

As in the *Collostructional Analysis* (cf. Stefanowitsch and Gries, [2003](http://www.linguistics.ucsb.edu/faculty/stgries/research/2003_AS-STG_Collostructions_IJCL.pdf)), `collex_fye()` uses one-tailed *Fisher-Yates Exact* test whose *p*-~FisherExact~value is log-transformed to the base of 10 to indicate the collostruction strength between the collocates and the node word (cf., e.g., Gries, Hampe, and Schönefeld, [2005](http://www.linguistics.ucsb.edu/faculty/stgries/research/2005_STG-BH-DS_CollStr-vs-Freq_CogLing.pdf), *inter alia*). `collex_fye()` simultaneously performs two uni-directional measures of *Delta P*; one of these shows the extent to which the presence of the node-word cues the collocates/collexemes, and *vice versa*.

Here is the code to perform the SCA

```{r perform-fye}
# perform FYE test for Collexeme Analysis
am_fye <- collex_fye(df = assoc_tb, collstr_digit = 3)
```

Now we can retrieve the top-10 most strongly attracted collocates to *mengatakan* 'to say sth.'. The association strength is shown in the `collstr` column, which stands for *collostruction strength*. The higher, the stronger the association.

```{r attracted-collocates}
# get the top-10 most strongly attracted collocates
dplyr::top_n(am_fye, 10, collstr)
```

Column `a` contains the co-occurrence frequency of the collocates (`w`) with the `node` as its R1 collocates in the demo corpus. `p_fye` shows the one-tailed *p*~FisherExact~-value. Updated README file is prepared for retrieving data to perform *Distinctive Collexeme Analysis*.
